
      LOAD ALL CSV FILES CONVERT TO RDS

```{r}
setwd("C:/Users/Nishna/Documents/DMML_CW1/DMML_CW1_R/CW1/TASK5_6")
#Main dataset
dat <- dat <- read.csv("fer2018.csv", stringsAsFactors = FALSE)
saveRDS(dat,"fer2018.Rds")
rm(dat)

#happy
dat <- dat <- read.csv("fer2018happy.csv", stringsAsFactors = FALSE)
saveRDS(dat,"fer2018happy.Rds")
rm(dat)

#suprise
dat <- dat <- read.csv("fer2018surprise.csv", stringsAsFactors = FALSE)
saveRDS(dat,"fer2018surprise.Rds")
rm(dat)

#neutral
dat <- dat <- read.csv("fer2018neutral.csv", stringsAsFactors = FALSE)
saveRDS(dat,"fer2018neutral.Rds")
rm(dat)

#angry
dat <- dat <- read.csv("fer2018angry.csv", stringsAsFactors = FALSE)
saveRDS(dat,"fer2018angry.Rds")
rm(dat)

#sad
dat <- dat <- read.csv("fer2018sad.csv", stringsAsFactors = FALSE)
saveRDS(dat,"fer2018sad.Rds")
rm(dat)

#disgust
dat <- dat <- read.csv("fer2018disgust.csv", stringsAsFactors = FALSE)
saveRDS(dat,"fer2018disgust.Rds")
rm(dat)

#fear
dat <- dat <- read.csv("fer2018fear.csv", stringsAsFactors = FALSE)
saveRDS(dat,"fer2018fear.Rds")
rm(dat)

```

      SPLIT THE PIXEL COLUMNS OF ALL FILES

```{r}
setwd("C:/Users/Nishna/Documents/DMML_CW1/DMML_CW1_R/CW1/TASK5_6")
#main dataset
dat <- readRDS(file = "fer2018.Rds")
pixels_splitted <- cSplit(dat,"pixels",sep = " ")
saveRDS(pixels_splitted, "pixels_splitted.Rds")
rm(dat)


#happy
dat <- readRDS(file = "fer2018happy.Rds")
happy_pixels_splitted <- cSplit(dat,"pixels",sep = " ")
dim(happy_pixels_splitted)
saveRDS(happy_pixels_splitted, "happy_pixels_splitted.Rds")
rm(dat)

#surprise
dat <- readRDS(file = "fer2018surprise.Rds")
surprise_pixels_splitted <- cSplit(dat,"pixels",sep = " ")
dim(surprise_pixels_splitted)
saveRDS(surprise_pixels_splitted, "surprise_pixels_splitted.Rds")
rm(dat)

#neutral
dat <- readRDS(file = "fer2018neutral.Rds")
neutral_pixels_splitted <- cSplit(dat,"pixels",sep = " ")
dim(neutral_pixels_splitted)
saveRDS(neutral_pixels_splitted, "neutral_pixels_splitted.Rds")
rm(dat)

#angry
dat <- readRDS(file = "fer2018angry.Rds")
angry_pixels_splitted <- cSplit(dat,"pixels",sep = " ")
dim(angry_pixels_splitted)
saveRDS(angry_pixels_splitted, "angry_pixels_splitted.Rds")
rm(dat)

#sad
dat <- readRDS(file = "fer2018sad.Rds")
sad_pixels_splitted <- cSplit(dat,"pixels",sep = " ")
dim(sad_pixels_splitted)
saveRDS(sad_pixels_splitted, "sad_pixels_splitted.Rds")
rm(dat)

#disgust
dat <- readRDS(file = "fer2018disgust.Rds")
disgust_pixels_splitted <- cSplit(dat,"pixels",sep = " ")
dim(disgust_pixels_splitted)
saveRDS(disgust_pixels_splitted, "disgust_pixels_splitted.Rds")
rm(dat)

#fear
dat <- readRDS(file = "fer2018fear.Rds")
fear_pixels_splitted <- cSplit(dat,"pixels",sep = " ")
dim(fear_pixels_splitted)
saveRDS(fear_pixels_splitted, "fear_pixels_splitted.Rds")
rm(dat)
```


      LOAD THE MAIN DATASET  WITH PIXEL COLUMN SPLITTED
```{r}
setwd("C:/Users/Nishna/Documents/DMML_CW1/DMML_CW1_R/CW1")
pixels_splitted <- readRDS("pixels_splitted.Rds")
```
        
      PARTITION THE DATASET FOR TRAIN & TEST
```{r}
library(caret)
set.seed(2)
test_index <- createDataPartition(pixels_splitted$emotion,times = 1, p = 0.25, list = FALSE)
pixel_test <- pixels_splitted[test_index, ]
pixel_train <- pixels_splitted[-test_index, ]
```
        
      FREEZE TEST & TRAIN DATASET & SAVE THEM IN SEPERATE FILES 
```{r}
saveRDS(pixel_test, "pixel_test.Rds")
saveRDS(pixel_train, "pixel_train.Rds")

#Saving the predictor variables in matrix format in another Rds file
pixel_train_matrix <- as.matrix(pixel_train[,-1])
saveRDS(pixel_train_matrix, "pixel_train_matrix.Rds")
```
    
            LOADING ALL THE REQUIRED LIBRARIES
```{r}
library(caret)
library(dplyr)
library(matrixStats)
library(splitstackshape)
```


         LOADING THE TRAIN DATASET 
```{r}
setwd("C:/Users/Nishna/Documents/DMML_CW1/F21DL_CW1_R")
pixel_train <- readRDS("pixel_train.Rds")
pixel_train_matrix <- readRDS("pixel_train_matrix.Rds")
```

        VISUALIZE SOME IMAGES IN THE DATASET
```{r}
set.seed(123)
index <- sample(nrow(pixel_train_matrix), 25)
x <- pixel_train_matrix[index,]


library(EBImage)
grid <- matrix(x[1,], 48, 48,byrow = TRUE)
image(1:48, 1:48, grid) 

x_pixels_rotated <- apply(x,MARGIN = 1,FUN = function(x_row){
  grid <- matrix(x_row, 48, 48)
  #image(1:48,1:48,grid)
  img_rotate <- rotate(grid,90,bg.col = "white")
  as.vector(img_rotate)
  
  } )
x_pixels_rotated_90 <- rotate(x_pixels_rotated,90)

i=1
while(i < nrow(x_pixels_rotated_90))
{
  grid <- matrix(x_pixels_rotated_90[i,], 48, 48,byrow = TRUE)
  image(1:48, 1:48, grid) 
  i = i + 1
}



```



        VISUALIZE THE DISTRIBUTION & FILTER FEATURES BASED ON SUMMARY 
```{r}
library(matrixStats)
sds <- colSds(pixel_train_matrix)

grid <- matrix(sds, 48, 48,byrow = TRUE)
image(1:48, 1:48, grid,col= topo.colors(12)) 

qplot(sds, bins = "30", color = I("black"))
```
        

```{r}
rm_pixels <- pixel_train_matrix[ ,sds < 53.25]

#Finding attributes to be retained
col_index <- setdiff(1:ncol(pixel_train_matrix), 1:ncol(rm_pixels) )
print("No of features in the new dataset :") 
length(col_index)
new_pixel_train_matrix <- pixel_train_matrix[,col_index]
```    
    
          SCALE & PCA
```{r}
pixel_scaled <- scale(new_pixel_train_matrix,scale = TRUE, center = TRUE)
pca_scaled <- prcomp(pixel_scaled)
plot(pca_scaled$sdev)
saveRDS(pca_scaled, file = "pca_scaled.Rds")
library(dplyr)
summary(pca_scaled)$importance[,c(14,34,68,100,300)] %>% knitr::kable()
```
        
          TRAINING STRATIFIED SAMPLEs (1000 from each emotions )for k=3
```{r}
pca_scaled <- readRDS("pca_scaled.Rds")
set.seed(123)
Strata_index <- stratified(pixel_train, "emotion", size = 1000)
index_strata <- sample(which(Strata_index$emotion %in% pixel_train$emotion))
y <- pixel_train$emotion[index_strata,drop = FALSE]
control <- trainControl(method = "cv", number = 10, p = .9,
                        repeats = 5,verboseIter = TRUE)
```
```{r}
#PC 1 : 14
K <- 14
train_knn_14_scaled <- train(pca_scaled$x[index_strata, 1 : K], y,
                             method = "knn",
                             tuneGrid =  data.frame(k = 3),
                             trControl = control)
#PC 1 : 24
K <- 24
train_knn_24_scaled <- train(pca_scaled$x[index_strata, 1 : K], y,
                             method = "knn",
                             tuneGrid =  data.frame(k = 3),
                             trControl = control)

#PC 1 : 28
K <- 28
train_knn_28_scaled <- train(pca_scaled$x[index_strata, 1 : K], y,
                             method = "knn",
                             tuneGrid =  data.frame(k = 3),
                             trControl = control)

#PC 1 : 34
K <- 34
train_knn_34_scaled <- train(pca_scaled$x[index_strata, 1 : K], y,
                             method = "knn",
                             tuneGrid =  data.frame(k = 3),
                             trControl = control)

#PC 1 : 68
K <- 68
train_knn_68_scaled <- train(pca_scaled$x[index_strata, 1 : K], y,
                             method = "knn",
                             tuneGrid =  data.frame(k = 3),
                             trControl = control)

#PC 1 : 100
K <- 100
train_knn_100_scaled <- train(pca_scaled$x[index_strata, 1 : K], y,
                             method = "knn",
                             tuneGrid =  data.frame(k = 3),
                             trControl = control)
#PC 1 : 140
K <- 140
train_knn_140_scaled <- train(pca_scaled$x[index_strata, 1 : K], y,
                             method = "knn",
                             tuneGrid =  data.frame(k = 3),
                             trControl = control)



#PC 1 : 200
K <- 200
train_knn_200_scaled <- train(pca_scaled$x[index_strata, 1 : K], y,
                             method = "knn",
                             tuneGrid =  data.frame(k = 3),
                             trControl = control)


#PC 1 : 300
K <- 300
train_knn_300_scaled <- train(pca_scaled$x[index_strata, 1 : K], y,
                             method = "knn",
                             tuneGrid =  data.frame(k = 3),
                             trControl = control)

save(train_knn_14_scaled,
     train_knn_24_scaled,
     train_knn_28_scaled,
     train_knn_34_scaled,
     train_knn_68_scaled,
     train_knn_100_scaled,
     train_knn_200_scaled,
     train_knn_300_scaled, file = "knn_pca.rda")
```

      PLOT RMSE VS NO OF FEATURES
```{r}
load("knn_pca.rda")
#Plot the results
k <- c(14,24,28,34,68,100,140,200,300)
rmse_list <- c(train_knn_14_scaled$results$RMSE,
              train_knn_24_scaled$results$RMSE,
              train_knn_28_scaled$results$RMSE,
              train_knn_34_scaled$results$RMSE,
              train_knn_68_scaled$results$RMSE,
              train_knn_100_scaled$results$RMSE,
              train_knn_140_scaled$results$RMSE,
              train_knn_200_scaled$results$RMSE,
              train_knn_300_scaled$results$RMSE)

rmse_1000 <- data.frame(K = k, RMSE = rmse_list)

rmse_1000 %>% ggplot(aes(rmse_1000$K,rmse_1000$RMSE, col = K)) +
  geom_point(size = 2) + 
  xlab("Principal Components 1:K") + ylab("Root Mean Squared Error") +
  geom_line()


#Find out the model with min RMSE
K_win <- rmse_1000$K[which.min(rmse_1000$RMSE)]
K_win
```


```{r}
#Remove the train models for dealing with computational constraints
rm(train_knn_100_scaled)
rm(train_knn_14_scaled)
rm(train_knn_140_scaled)
rm(train_knn_200_scaled)
rm(train_knn_24_scaled)
rm(train_knn_28_scaled)
rm(train_knn_300_scaled)
rm(train_knn_34_scaled)
rm(train_knn_68_scaled)
rm(Strata_index)
rm(rm_pixels)
rm(pixel_scaled)
rm(new_pixel_train_matrix)
rm(control)
rm(index_strata)
rm(rmse_list)
rm(rmse_1000_scaled)
rm(k)
rm(sds)
rm(K)
rm(train_knn_10)
rm(train_knn_2)
rm(train_knn_5)
K_win
```

```{r}
#fit the winner model to the whole dataset with k = 3
y <- factor(pixel_train$emotion)
control <- trainControl(method = "cv", number = 10, p = .9,
                        verboseIter = TRUE)
fit_K_win <- train(pca_scaled$x[, 1 : K_win], y,
                   method = "knn",
                   tuneGrid =  data.frame(k = 3),
                   trControl = control)

```

      COMPUTE TEST ACCURACY
        
```{r}
setwd("C:/Users/Nishna/Documents/DMML_CW1/F21DL_CW1_R")
pixel_test <- readRDS("pixel_test.Rds")
dim(pixel_test)

#convert to matrix
pixel_test_matrix <- as.matrix(pixel_test[,-1])
dim(pixel_test_matrix)

#filter features
pixel_test_matrix <- pixel_test_matrix[,col_index]
dim(pixel_test_matrix)

#Transform the whole test data
col_means <- colMeans(pixel_test_matrix)
x_test <- sweep(pixel_test_matrix, 2, col_means) %*% pca_scaled$rotation
x_test <- x_test[,1 :K_win]

#Prediction
y_hat <- predict(fit_K_win, x_test, type = "raw")

#Results
confusionMatrix(y_hat, factor(pixel_test$emotion))
```
                   
                   FIT THE MODELS TO THE FULL TRAINING DATASET
            
```{r}
#fit PC 1: 14 to full data
fit_K_14 <- train(pca_scaled$x[, 1 : 14], y,
                   method = "knn",
                   tuneGrid =  data.frame(k = 3),
                   trControl = control)

#fit PC 1: 34 to full data
fit_K_34 <- train(pca_scaled$x[, 1 : 34], y,
                   method = "knn",
                   tuneGrid =  data.frame(k = 3),
                   trControl = control)

#fit PC 1: 68 to full data
fit_K_68 <- train(pca_scaled$x[, 1 : 68], y,
                   method = "knn",
                   tuneGrid =  data.frame(k = 3),
                   trControl = control)
save(fit_K_14, fit_K_34, fit_K_68, file = "pca_models_14_34_68.rda")

```
            
            